---
title: "Airline Passenger Satisfaction"
author: "Basharat Asghar"
date: '2022-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

#### Required Libraries

```{r}
library(ggplot2)
library(vroom)
library(dplyr)
library(forcats)
library(tidyr)
library(skimr)
library(janitor)
```

#### Reading data

```{r}
passenger_df <- vroom("airline_passenger_satisfaction.csv")
```

#### Look into data frame structure and statistics

```{r}
skim(passenger_df)
```

Data frame consists of more than 100K observations with 24 columns. Data frame have 5 character variables {columns}, they have no missing value.14 columns consist of numeric type, only Arrival Delay have missing values, for most of the variables distributions is not skewed.

#### Converting names to standard r names with janitor

```{r}
passenger_df <- passenger_df %>% 
  clean_names()
```

```{r,results='hide',include=FALSE,eval=FALSE}
passenger_df %>% 
  get_dupes()
```

#### Removing values containing 'NA'

```{r}
passenger_df <- passenger_df %>% 
  filter(!is.na(arrival_delay))
```

### Exploratory Data Analysis {EDA}

#### Numerical Data

**Converting Age into Age-Categories**  

```{r}
passengers_df1 <- passenger_df %>% 
  mutate(age = case_when(age >= 0 & age <=12 ~ "Children",
                         age >=13 & age <=19 ~ "Teenagers",
                         age >=20 & age <=25 ~ "Youngsters",
                         age >=26 & age <=55 ~ "MiddleAge",
                         age > 55  ~ "Old"),
         age = as.factor(age)) %>% 
  mutate(flight_distance = case_when(flight_distance > 0 & flight_distance <= 500 ~ "short-distance",
                                     flight_distance > 500 & flight_distance <= 1500 ~ "Midrange-distance",
                                     flight_distance > 1500 & flight_distance <= 3000 ~ "MidrangePlus-distance",
                                     flight_distance > 3000 & flight_distance <= 5000 ~ "Long-distance"),
         flight_distance = as.factor(flight_distance)) %>% 
  mutate(departure_delay = case_when(departure_delay == 0 ~ "No-delay",
                                     departure_delay > 0 & departure_delay <= 100 ~ "0-100",
                                     departure_delay > 100 & departure_delay <= 300 ~ "100-300",
                                     departure_delay > 300 & departure_delay <= 500 ~ "300-500",
                                     departure_delay > 500 ~ "> 500"),
         departure_delay = as.factor(departure_delay)) %>% 
  mutate(arrival_delay = case_when(arrival_delay == 0 ~ "No-delay",
                                   arrival_delay > 0 & arrival_delay <= 100 ~ "0-100",
                                   arrival_delay > 100 & arrival_delay <= 300 ~ "100-300",
                                   arrival_delay > 300 & arrival_delay <= 500 ~ "300-500",
                                   arrival_delay > 500 ~ "> 500"),
         arrival_delay = as.factor(arrival_delay))
```

```{r}
passengers_df1 %>% 
  ggplot(aes(age, fill = age)) +
  geom_bar(position = "dodge", show.legend = FALSE) +
  facet_wrap(vars(satisfaction), ncol = 2) +
  scale_fill_manual(values = c(MiddleAge = "#219ebc", "darkgrey", "darkgrey", "darkgrey",
                               "darkgrey")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  labs(x = "Age",
       y = "Count of Passengers",
       title = "Passengers who are satisfied or not w.r.t Age Categories",
       caption = "Data Source : Maven Analytics") +
  coord_flip()
```

Middle age passengers are about 62% of the total, and little over 50% are dis-satisfied with the airline. For different genders trend for dis-satisfied or satisfied customers is the same.\
After Middle age, Old and young customers make about 29% of the passengers, and from these 61% are dis-satisfied.

```{r}
passengers_df1 %>% 
  ggplot(aes(flight_distance, fill = satisfaction)) +
  geom_bar(position = "dodge", show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Flight Distance (Miles)",
       y = "Count of Passengers",
       title = "Passengers who are satisfied or not w.r.t Flight Distance",
       caption = "Data Source : Maven Analytics") +
  coord_flip()
```

Mid-Range & Short-Range distances categories have about 70% customers, and about 46% from them are dis-satisfied.  

```{r,warning=FALSE}
passengers_df1 %>%
  ggplot(aes(departure_delay, fill = satisfaction)) +
  geom_bar(position = "dodge", show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  scale_x_discrete(limits = c(">500", "300-500", "100-300", "0-100", "No-delay")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Departure Delay (Minutes)",
       y = "Count of Passengers",
       title = "Passengers who are satisfied or not w.r.t Departure Delay",
       caption = "Data Source : Maven Analytics") +
  coord_flip()
```

Total 56-57% are dis-satisfied, and 0-100 & no-delay categories have up-to 97% of passengers from these two we have about 54-55% dis-satisfied customers.

```{r}
passengers_df1 %>%
  ggplot(aes(arrival_delay, fill = satisfaction)) +
  geom_bar(position = "dodge", show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  scale_x_discrete(limits = c(">500", "300-500", "100-300", "0-100", "No-delay")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Arrival Delay (Minutes)",
       y = "Count of Passengers",
       title = "Passengers who are satisfied or not w.r.t Arrival Delay",
       caption = "Data Source : Maven Analytics") +
  coord_flip()
```

Same is true for arrival delay as 57% are dis-satisfied & majority are from two mentioned categories.

```{r}
passenger_df %>%
  mutate(departure_and_arrival_time_convenience =     as_factor(departure_and_arrival_time_convenience)) %>% 
  ggplot(aes(departure_and_arrival_time_convenience, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Departure & Arrival time convience",
       y = "Count of passengers",
       title = "Time convience w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(ease_of_online_booking = as_factor(ease_of_online_booking)) %>%
  ggplot(aes(ease_of_online_booking, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Ease of online booking",
       y = "Count of passengers",
       title = "Ease of online booking w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(check_in_service = as_factor(check_in_service)) %>% 
  ggplot(aes(check_in_service, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Check-In Service",
       y = "Count of passengers",
       title = "Check-In Service w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(online_boarding = as_factor(online_boarding)) %>% 
  ggplot(aes(online_boarding, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Online Boarding",
       y = "Count of passengers",
       title = "Online Boarding w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(gate_location = as_factor(gate_location)) %>% 
  ggplot(aes(gate_location, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Gate Location",
       y = "Count of passengers",
       title = "Gate location w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(on_board_service = as_factor(on_board_service)) %>% 
  ggplot(aes(on_board_service, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "On-Board Service",
       y = "Count of passengers",
       title = "On-Board Service w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(seat_comfort = as_factor(seat_comfort)) %>% 
  ggplot(aes(seat_comfort, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Seat Comfort",
       y = "Count of passengers",
       title = "Seat Comfort w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(leg_room_service = as_factor(leg_room_service)) %>% 
  ggplot(aes(leg_room_service, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Leg Room Service",
       y = "Count of passengers",
       title = "Leg Room Service w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(cleanliness = as_factor(cleanliness)) %>% 
  ggplot(aes(cleanliness, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Cleanliness",
       y = "Count of passengers",
       title = "Cleanliness w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(food_and_drink = as_factor(food_and_drink)) %>% 
  ggplot(aes(food_and_drink, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Food and Drink",
       y = "Count of passengers",
       title = "Food and Drink w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(in_flight_service = as_factor(in_flight_service)) %>% 
  ggplot(aes(in_flight_service, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "In-Flight Service",
       y = "Count of passengers",
       title = "In-Flight Service w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(in_flight_wifi_service = as_factor(in_flight_wifi_service)) %>% 
  ggplot(aes(in_flight_wifi_service, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "In-Flight WiFi Service",
       y = "Count of passengers",
       title = "In-Flight WiFi Service w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(in_flight_entertainment = as_factor(in_flight_entertainment)) %>% 
  ggplot(aes(in_flight_entertainment, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "In-Flight Entertainment",
       y = "Count of passengers",
       title = "In-Flight Entertainment w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>%
  mutate(baggage_handling = as_factor(baggage_handling)) %>% 
  ggplot(aes(baggage_handling, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Baggage Handling",
       y = "Count of passengers",
       title = "Baggage Handling w.r.t Satisfaction",
       caption = "Data Source : Maven Analytics")
```

```{r}
avg_ratings <- passenger_df %>%
  summarise(across(departure_and_arrival_time_convenience:baggage_handling, mean, na.rm=TRUE)) %>% 
  pivot_longer(departure_and_arrival_time_convenience:baggage_handling,
               names_to = "service", values_to = "avg_rating") %>%
  mutate(service = as.factor(service),
         service = fct_reorder(service, avg_rating))

avg_ratings %>% 
  ggplot(aes(service, avg_rating, fill = service)) +
  geom_col(show.legend = FALSE) +
  geom_hline(yintercept = mean(avg_ratings$avg_rating), color = "#ee9b00", size = 1.5) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.subtitle = element_text(color = "darkgrey", face = "bold", size = 9),
        plot.caption = element_text(color = "darkgrey", face = "bold")) +
  labs(x = "Service Provided",
       y = "Average rating for each Service",
       title = "Average rating for each Service provided by Air-line",
       subtitle = "Line represents the average of ratings for all services",
       caption = "Maven Analytics") +
  scale_fill_manual(values = c(food_and_drink = "darkred",  departure_and_arrival_time_convenience = "darkred", gate_location = "darkred",
                               ease_of_online_booking = "darkred",
                               in_flight_wifi_service = "darkred", "darkgrey",
                               "darkgrey", "darkgrey", "darkgrey", "darkgrey",
                               "darkgrey", "darkgrey", "darkgrey", "darkgrey")) +
  coord_flip()
```

In-Flight-WiFi-Service, Ease-of-Online-Booking, Gate-Location have the lowest average for ratings.\
In-Flight-Service have the highest average rating but about 62% are the dis-satisfied customers, as compare to other services where dis-satisfied customers are in the range of 55-57%.

### Categorical Data

```{r}
passenger_df %>% 
  ggplot(aes(satisfaction, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.subtitle = element_text(color = "darkgrey", face = "bold", size = 9),
        plot.caption = element_text(color = "darkgrey", face = "bold")) +
  labs(x = "Satisfaction",
       y = "Count of Passengers",
       title = "No.of customers who are satisfied or not",
       caption = "Data Source : Maven Analytics") +
  scale_fill_manual(values = c("#1d3557", "#457b9d"))
```

**57% customers are dis-satisfied over-all.**  

```{r}
passenger_df %>% 
  ggplot(aes(gender, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Gender",
       y = "Count of Passengers",
       title = "No.of customers who are satisfied or not for Gender",
       caption = "Data Source : Maven Analytics")
```

```{r}
passenger_df %>% 
  ggplot(aes(customer_type, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Type of customers",
       y = "Count of Passengers",
       title = "No.of customers who are satisfied or not for Type of Customers",
       caption = "Data Source : Maven Analytics")
```

First-Time customers are 18% of the total but 76% are dis-satisfied.\
Returning are more in numbers and 52% are dis-satisfied.

```{r}
passenger_df %>% 
  ggplot(aes(type_of_travel, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Travel Type",
       y = "Count of Passengers",
       title = "No.of customers who are satisfied or not for Travel Type",
       caption = "Data Source : Maven Analytics")
```

Personal travelers are about 31% but 90% from these are dis-satisfied.

```{r}
passenger_df %>% 
  ggplot(aes(class, fill = satisfaction)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("#1d3557", "#457b9d")) +
  theme_light() +
  theme(axis.title.x = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.x = element_text(face = "bold", size = 7, color = "darkgrey"),
        axis.title.y = element_text(face = "bold",size = 9, color = "darkgrey"),
        axis.text.y = element_text(face = "bold", size = 7, color = "darkgrey"),
        plot.title = element_text(color = "darkgrey", face = "bold", size = 11),
        plot.caption = element_text(color = "darkgrey", face = "bold"),
        strip.text.x = element_text(face = "bold", size = 8, color = "darkgrey"),
        strip.background = element_rect(fill = "white")) +
  facet_wrap(vars(satisfaction)) +
  labs(x = "Travel Class",
       y = "Count of Passengers",
       title = "No.of customers who are satisfied or not for Travel Class",
       caption = "Data Source : Maven Analytics")
```

Business class is doing good with almost 70% satisfied customers.\
Economy is having worst customer experience as they count for 45% of total passengers and more than 80% are dis-satisfied.\
Economy Plus have lowest count of customers but there 75% customers are also dis-satisfied.

### Key Takeaways:

1.  Middle age passengers are about 62% of the total, and little over 50% are dis-satisfied with the airline. For different genders trend for dis-satisfied or satisfied customers is the same.\
    After Middle age, Old and young customers make about 29% of the passengers, and from these 61% are dis-satisfied.

2.  Mid-Range & Short-Range distances categories have about 70% customers, and about 46% from them are dis-satisfied.

3.  0-100 & no-delay categories have up-to 97% of passengers from these two we have about 54-55% dis-satisfied customers.

4.  Same is true for arrival Delay.

5.  In-Flight-WiFi-Service, Ease-of-Online-Booking, Gate-Location, Departure_arrival_time_convenience, Food_and_Drink have the average rating less than overall average rating.

6.  In-Flight-Service have the highest average rating but about 62% are the dis-satisfied customers, as compare to other services where dis-satisfied customers are in the range of 55-57%.

7.  Overall 57% customers are dis-satisfied.

8.  Female are 57% and males are 56% dis-satisfied.

9.  First-Time customers are 18% of the total but 76% are dis-satisfied.\
    Returning are more in numbers and 52% are dis-satisfied.

10. Personal travelers are about 31% but 90% from these are dis-satisfied.

11. Business class is doing good with almost 70% satisfied customers.\
    Economy is having worst customer experience as they count for 45% of total passengers and more than 80% are dis-satisfied.\
    Economy Plus have lowest count of customers but there 75% customers are also dis-satisfied.  
    
### Recommendations :  
1. Middle-Age customers should get more attention, as they are in bulk.  
   - Young & Old customers must also get good services.  
2. In-Flight-WiFi-Service, Ease-of-Online-Booking, Gate-Location, Food_and_Drink,
   Departure_arrival_time_convenience, these services should be improved on, as they have
   the average rating lower than overall average rating.  
3. In-Flight-Service have gets highest average rating, but still need improvements, as 
   most dis-satisfied percentage of customers is also from this service.  
4. First-Time customers should get high quality service, so, that they can be convinced
   to return more often.  
5. Personal travelers & Economy & Economy Plus class travelers should also receive high
   quality service.  
6. Overall all services provided by company need improvements.  
